---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)


av <- av %>% rowwise() %>% mutate(Death_Time = sum(c(Death1, Death2, Death3, Death4, Death5) == "YES"))
av <- av %>% rowwise() %>% mutate(Death = Death_Time > 0) %>% select(-c(Death1, Death2, Death3, Death4, Death5))
av$Death <- ifelse(av$Death == TRUE, "yes", "no")
av
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.
```{r}
av <- av %>% rowwise() %>% mutate(Returns_Time = sum(c(Return1, Return2, Return3, Return4, Return5) == "YES"))
av <- av %>% rowwise() %>% mutate(Returns = Returns_Time > 0) %>% select(-c(Return1, Return2, Return3, Return4, Return5))
av$Returns <- ifelse(av$Returns == TRUE, "yes", "no")
av

```
Based on these datasets calculate the average number of deaths an Avenger suffers. 
The average number of avenger deaths is 0.3294798.
```{r}
av
mean(av$Deaths)

```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement
Grace: I counted 89 total deaths — some unlucky Avengers7 are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback. 

```{r}
sum(av$Death_Time)
sum(av$Returns_Time)
```
The code shows that the total number of times the avengers have died is indeed 89. The article is also correct claiming that there are 57 occasions of the avenger making a comeback.


> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.


##Blake
For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.

### Include the code

```{r}
avengers.with.death = nrow(av[av$Death == "yes",])
avengers.with.death
```

### Include your answer

We can see that the number of rows with at least one death is the same as the number that the author found in his analysis.

## Grace

### FiveThirtyEight Statement

> I counted 89 total deaths — some unlucky Avengers7 are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback. 

```{r}
sum(av$Death_Time)
sum(av$Returns_Time)
```

The code shows that the total number of times the avengers have died is indeed 89. The article is also correct claiming that there are 57 occasions of the avenger making a comeback.


## Nick

### FiveThirtyEight Statement

> There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death.

### Included code

```{r}
# number of 1st returns divided by number of 2nd deaths
 sum(nrow(av[av$Returns_Time >= 1, "Returns_Time"])) / sum(nrow(av[av$Death_Time >= 1, "Death_Time"]))

# number of 2nd returns divided by number of 2nd deaths
 sum(nrow(av[av$Returns_Time >= 2, "Returns_Time"])) / sum(nrow(av[av$Death_Time >= 2, "Death_Time"]))
 
# number of 3rd returns divided by number of 3rd deaths
 sum(nrow(av[av$Returns_Time >= 3, "Returns_Time"])) / sum(nrow(av[av$Death_Time >= 3, "Death_Time"]))
 
# number of 3rd deaths
 nrow(av[av$Death_Time >= 3, "Death_Time"])

```

### Our Analysis:

We find that the author was correct on all counts, though there are only two characters which have died 3 times and one of them made a comeback (Jocasta).
